*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-04 Code/Manuscript_O2.Rmd
minor edits

We used the OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) [@saitoDevelopmentOceanProtein2021] to analyze the distribution of proteins from clades of *P. marinus* in samples taken across a range of [O~2~] and depth, which in turn correlates to peak light at the site of sampling. In parallel we analyzed the growth and physiological responses of representative strains from three clades of *P. marinus* under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods, to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *Prochlorococcus marinus* MED4, a clade HLI strain, was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *Prochlorococcus marinus* SS120, a clade LLIII strain, was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV strain, was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue light, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

@@

Three xenic cultures of *P. marinus* were obtained from Bigelow Labs, NCMA Maine, USA. MED4 (CCMP1986) is from High-Light adapted (HLI) clade; SS120 (CCMP1375) is from Low-Light adapted (LLIII) clade; and MIT9313 (CCMP2773) is from Low-Light adapted (LLIV) clade. Cultures were maintained in incubators set to 22°C with a light/dark cycle of 12 h. The PAR level for maintenance cultures reflected PAR in the source niche of the ecotype; MED4, of 160 µmol photons m^-2^ s^-1^ with illumination from STANDARD Products Inc. Cool White F24T5/41K/8/HO/PS/G5/STD, 24 watts, fluorescent bulbs; SS120 and MIT9313 at 30 µmol photons m^-2^ s^-1^ with illumination from Philips Cool White F14T5/841 Alto, 14 watts, fluorescent bulbs. To maintain active growth all strains were transferred weekly with 1 in 5 dilutions with Pro99 media [@mooreCulturingMarineCyanobacterium2007] prepared with autoclaved artificial seawater [@bergesEvolutionArtificialSeawater2001].

@@

## Growth rate analysis {.unnumbered}
Data files (.csv) saved from the Multicultivator software were imported into R-Studio for data management [@R-tidyverse], growth rate calculations CITATION, comparisons of model fits [@harrisonPhotosynthesisirradianceRelationshipsPolar1986], and visualization. The chlorophyll proxy optical density (OD~680~ - OD~720~;  ΔOD) was used to determine the chlorophyll specific growth rate (µ, d^-1^) for each treatment combination.  We first used a rolling mean from the R package *`zoo`* [@R-zoo] to calculate the average ΔOD data over a 1-hour window to lower the influence of outlier points and remove data points collected during post stationary phase, when applicable. We used the Levenberg-Marquardt algorithm [@bellaviaLevenbergMarquardtMethod2018] modification of the non-linear least squares,  using the R package *`minpack.lm`*
[@R-minpack.lm], to fit a logistic equation (Equation \@ref(eq:GrowthRate)); where ΔOD~max~ is maximum ΔOD, ΔOD~min~ is minimum ΔOD, t is time
duration over the growth trajectory.

@@

XXX Some text is relevant when referring to original NATLProchlorPhotosynDepthO2 figure, not with this figureXXX
Proteins from 13 annotated strains of *P. marinus* were detected across depths and oxygen concentrations in the ocean proteins data set analyzed. We focused our analysis here on photosynthetic protein complexes, grouped by clade. Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) plots the observation of core photosynthetic complexes (Photosystem II, the Cytochromeb6f complex, Phototsystem I, ATP Synthase and Rubisco) across clades, as a function of depth (a proxy for light intensity) and measured [O~2~]. Photosynthetic complexes from HLI (ex. MED4) were detected throughout the water column, predominantly at high [O~2~]. Complexes from Clade LLIII (ex. SS120) were detected throughout the water column in high [O~2~] and at depth in low [O~2~] samples. Photosystem I and II were only detected at low [O~2~] XXX This is evident when referring to original NATLProchlorPhotosynDepthO2 figure, not with this figureXXX. The exclusive assignment of ATP Synthase peptides to Clade IV (ex. MIT9313) may result from the high conservation of protein sequence for this complex.

@@

CladeProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (clade HLI), SS120 (clade LLIII) and MIT9313 (clade LLIV).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

*P. marinus* SS120 clade LLIII,  growth under 250 µM O~2~ increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and across all spectral wavebands (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^ but showed growth inhibition at 16 h red light photoperiod.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white LED, red or blue light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

@@

SS120PAR_cap <- glue("**chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 (Low-Light (LLIII) deep ocean clade) vs. photoperiod (h). **",
                     "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "colours represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

XXXXCHECK figure references here; not sure White Light data is shown on the bluered figure; refer to Supplemental?XXXX

@@

### SS120 clade LLIII {.unnumbered}

The representative of the LLIII clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments (Figure \@ref(fig:PhotoperiodPurFitsPlots)B). Most 4 h photoperiod treatments of SS120 also did not grow under 250 µM O~2~, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments (Figure \@ref(fig:PhotoperiodPurFitsPlots)B). SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments (Figure \@ref(fig:PhotoperiodPurFitsPlots)B).

@@

\newpage
XXXXEdit text depending upon which ocean protein figures we useXXXXX
Ocean detections of proteins mediating protein metabolism support this interpretation of distinct FtsH function across clades of *P. marinus*. Ribosome proteins from clade HLI MED4, clade LLIII SS120 and clade LLIV MIT9313 show generally similar patterns vs. [O~2~] and depth, a proxy for peak PAR  (Figure \@ref(fig:NATLProchloroProteinMetabDepthO2)).  FtsH3, inferred to mediate PSI assembly, likewise shows a similar pattern between MED4 and MIT9313 (Figure \@ref(fig:NATLProchloroProteinMetabDepthO2)).  But only MED4 shows detected presence of the FtsH1 & FtsH2 isoforms inferred to mediate PSII repair, and then only in near-surface samples subject to higher light.  Furthermore, even though MIT9313 grows (Figure \@ref(fig:MIT9313GAM)), and is detected in the ocean at low [O~2~] (Figure \@ref(fig:CladeProchloroPhotosynthDepthO2)), no FtsH from MIT9313 is detected at low [O~2~] (Figure \@ref(fig:NATLProchloroProteinMetabDepthO2)), suggesting limited requirement for protein turnover under low [O~2~].

@@

NATLProchloroProteinMetabDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* protein metabolism complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (clade HLI), SS120 (clade LLIII) and MIT9313 (clade LLIV).",
                       "Columns show detections of proteins annotated as FtsH Protease Complexes (FtsH1, FtsH2, FtsH3) or the Ribosome).",
                       "Culture growth experimental conditions indicated by horizontal grey lines for depths approximating Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^) and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

Figure \@ref(fig:ProchlorococcusEnzymeKms) shows the measured or inferred K~M~ for [O~2~] for genes encoding enzymes [@omarAnnotationGenesEncoding2023] for  *P. marinus* strains from clades HLI, LLI, LLIII and LLIV. XXXX CITATION  to Table once added XXXX. MED4 increases expression of alternative oxidase ('ubiquinol oxidase (non electrogenic)')  to cope with changes in light [@bergResponsesPsbAHli2011], by dissipating electrons from the inter-system transport chain. The approximate K~M~ for [O~2~] of ~ 25 µM for ubiquinol oxidase (non electrogenic) (Figure \@ref(fig:ProchlorococcusEnzymeKms)) is comparable to the lower limit for growth of MED4 in our experiments (Figure \@ref(fig:MED4GAM)).  We suggest that dependence upon this enzyme excludes MED4 from low oxygen zones.  The genome scan shows SS120 and MIT9313 lack this gene (Figure \@ref(fig:ProchlorococcusEnzymeKms)), and therefore, lack this oxygen-dependent path to cope with changing excitation.  Conversely, a gene encoding (S)-2-hydroxy-acid oxidase is encoded in the MIT9313 genome (Figure \@ref(fig:ProchlorococcusEnzymeKms)).  (S)-2-hydroxy-acid oxidase catalyzes the reaction of 2-hydroxy acid with O~2~ to produce toxic H~2~O~2~ [@cunaneCrystalStructureAnalysis2005]. (S)-2-hydroxy-acid oxidase has an approximate K~M~ for [O~2~] of ~ 250 µM, and produces H~2~O~2~, so growth at lower [O~2~] may protect MIT9313 from auto-intoxication from production of H~2~O~2~. We hypothesize that under 250 µM O~2~ and higher blue light, *P. marinus* MIT9313 suffered photoinhibition, resulting from the inactivation of PSII caused by the production of the reactive oxygen species, hydrogen peroxide. This photoinhibition is compounded by the limited inducible repair mechanism for PSII, due to the absence of FtsH 1 and 2 XXXXCHECK TERMINOLOGYXXXX expression in *P. marinus* MIT9313 [@bonisteelStrainSpecificDifferences2018].  We hypothesize that under the conditions of our high light and 2.5 µM or 25 µM O~2~ experiments, the activity of the (S)-2-hydroxy-acid oxidase enzyme is suppressed. As a result, the catalyzed production of hydrogen peroxide is inhibited, leading to less PSII damage, allowing MIT9313 to avoid photoinhibition and circumvent its limitations on PSII repair to exploit higher light.  Figure \@ref(fig:ProchlorococcusEnzymeKms) also shows that *P. marinus* SS120 is the only tested ecotype to lack the pyridoxal 5'-phosphate synthase enzyme. The pyridoxal 5'-phosphate synthase enzyme is an important cofactor in the biosynthesis of vitamin B~6~ [@sangIdentificationPyridoxinePyridoxamine2007]. Vitamin B~6~ is a potential antioxidant and can effectively quench singlet oxygen [@bilskiVitaminB6Pyridoxine2000]. The absence of the pyridoxal 5'-phosphate synthase enzyme may explain why *P. marinus* SS120 does not grow as well as *P. marinus* MIT9313, when exposed to high light stress under 25 µM O~2~ and not at all under 2.5 µM O~2~ (Figure \@ref(fig:SS120PAR)).

@@

DNARepairFig_cap <- glue("**Genes encoding DNA repair enzymes.**" ,
                         "The y-axis represents  *Prochlorococcus marinus* strains.",
                         "The x-axis represents enzymes encoded for DNA repair found in at least one *Prochlorococcus marinus* strain in this study.",
                         "The red solid circles represent a gene count of 1.",
                         "The green solid circles represent a gene count of 2.",
                         "The blue solid circles represent a gene count of 3.",
                       "Figure was generated using a filtered subset of the annotated phytoplankton gene sequences dataset from Omar *et al*. [@omarAnnotationGenesEncoding2023].",

@@

We analyzed  growth rates to determine the viability of *P. marinus* MED4, a clade HLI ecotype found at depths near the ocean surface,
*P. marinus* SS120, a clade LLIII ecotype found deep in the water column, and *P. marinus* MIT9313, a clade LLIV ecotype also found in deep oceans, under present day and hypothetical future niche conditions. We designed our study to account for light attenuation through ocean waters, recognizing that low light levels of the blue spectral waveband ultimately prevail in deep ocean waters [@morelAvailableUsableStored1978; @holtropVibrationalModesWater2021] and that the solar incidence angle also attenuates the photoperiod or hours of light exposure beneath the ocean surface. We also tested the growth response of *P. marinus* to varying [O~2~] under the matrix of light parameters to determine if *P. marinus* strains can cope with  diverse  ecosystem scenarios.

@@

*Prochlorococcus marinus* SS120, a LLIII clade representative, showed an interactive inhibition of growth by oxygen and cumulative diel PUR, with a higher tolerance for higher cumulative diel PUR under 25 µM O~2~, compared to 250 µM O~2~. Thus, SS120 can exploit habitats with O~2~ levels spanning 25 µM O~2~ to 250 µM O~2~, including high PAR environments within OMZ.  SS120 is likely excluded from higher [O~2~] and higher PAR because of genomic limitations on capacity for DNA repair (Figure \@ref(fig:DNARepairFig)), and possibly by limited capacity for synthesis of reactive oxygen quenchers (Figure  \@ref(fig:ProchlorococcusEnzymeKms)).  Our growth results are supported by Lavin *et al*. [@lavinNovelLineagesProchlorococcus2010] who found evidence of LLII/III
and LLIV ecotypes, using terminal restriction fragment length polymorphism analyses, at depths above 40 m, where light levels are
higher, within OMZ, and by Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) showing PSII protein subunits annotated as derived from SS120 at all depths ranging
from 20 to 200 m and all [O~2~] in an OMZ of the tropical North Pacific Ocean. SS120, grew under photoperiods longer than 4 h and showed increasing growth rate with increasing photoperiods. Therefore, SS120 can potentially exploit the longer summer photoperiods in temperate zones however, like MED4, cannot sustain growth in winter through its requirement for more than 4 h of light per day, even if water temperature warms into the clade LLIII tolerance range.

*Prochlorococcus marinus* SS120 prefers the deeper ocean where blue spectral waveband dominates at low levels, through its demonstrated ability to exploit only the lowest blue spectral waveband levels PAR of 30 µmol photons m^-2^ s^-1^ under 250 µM O~2~. However, we found *P. marinus* SS120 can exploit more diverse ecological niches within the ocean layers, even in regions with high levels of blue spectral waveband,  but only under O~2~ of 25 µM (Figure \@ref(fig:SS120PAR)). Lavin *et al*. [@lavinNovelLineagesProchlorococcus2010] show evidence of LLIII ecotypes at varying depths from 20 to 30 m and from 75 to 200 µM O~2~ in the oxygen minimum zone (OMZ) of the tropical South Pacific Ocean, illustrating its tolerance of high light levels under low O~2~. SS120 has the potential to thrive in deep temperate zones, specifically during the spring, summer, and fall seasons when the duration of daylight exceeds 4 h, if O~2~ are near surface saturation of about 250 µM. Under lower oxygen levels of 25 µM, SS120 can also exploit a 4 h photoperiod in the blue waveband, and thus has the potential to inhabit a warmed, deep, temperate region during the winter season.

@@

>>>>

We used the OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) [@saitoDevelopmentOceanProtein2021] to analyze the distribution of proteins from clades of *P. marinus* in samples taken across a range of [O~2~] and depth, which in turn correlates to peak light at the site of sampling. In parallel we analyzed the growth and physiological responses of representative strains from three clades of *P. marinus* under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods, to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *Prochlorococcus marinus* MED4, a clade HLI strain, was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *Prochlorococcus marinus* SS120, a clade LLII/III strain, was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV strain, was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue light, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

@@

Three xenic cultures of *P. marinus* were obtained from Bigelow Labs, NCMA Maine, USA. MED4 (CCMP1986) is from High-Light adapted (HLI) clade; SS120 (CCMP1375) is from Low-Light adapted (LLII/III) clade; and MIT9313 (CCMP2773) is from Low-Light adapted (LLIV) clade. Cultures were maintained in incubators set to 22°C with a light/dark cycle of 12 h. The PAR level for maintenance cultures reflected PAR in the source niche of the ecotype; MED4, of 160 µmol photons m^-2^ s^-1^ with illumination from STANDARD Products Inc. Cool White F24T5/41K/8/HO/PS/G5/STD, 24 watts, fluorescent bulbs; SS120 and MIT9313 at 30 µmol photons m^-2^ s^-1^ with illumination from Philips Cool White F14T5/841 Alto, 14 watts, fluorescent bulbs. To maintain active growth all strains were transferred weekly with 1 in 5 dilutions with Pro99 media [@mooreCulturingMarineCyanobacterium2007] prepared with autoclaved artificial seawater [@bergesEvolutionArtificialSeawater2001].

@@

## Growth rate analysis {.unnumbered}
Data files (.csv) saved from the Multicultivator software were imported into R-Studio for data management [@R-tidyverse], growth rate calculations, comparisons of model fits [@harrisonPhotosynthesisirradianceRelationshipsPolar1986], and visualization. The chlorophyll proxy optical density (OD~680~ - OD~720~;  ΔOD) was used to determine the chlorophyll specific growth rate (µ, d^-1^) for each treatment combination.  We first used a rolling mean from the R package *`zoo`* [@R-zoo] to calculate the average ΔOD data over a 1-hour window to lower the influence of outlier points and remove data points collected during post stationary phase, when applicable. We used the Levenberg-Marquardt algorithm [@bellaviaLevenbergMarquardtMethod2018] modification of the non-linear least squares,  using the R package *`minpack.lm`*
[@R-minpack.lm], to fit a logistic equation (Equation \@ref(eq:GrowthRate)); where ΔOD~max~ is maximum ΔOD, ΔOD~min~ is minimum ΔOD, t is time
duration over the growth trajectory.

@@

Proteins from 13 annotated strains of *P. marinus* were detected across depths and oxygen concentrations in the ocean proteins data set analyzed. We focused our analysis here on photosynthetic protein complexes, grouped by clade. Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) plots the observation of core photosynthetic complexes (Photosystem II, the Cytochromeb6f complex, Phototsystem I, ATP Synthase and Rubisco) across clades, as a function of depth (a proxy for light intensity) and measured [O~2~]. Photosynthetic complexes from HLI (ex. MED4) were detected throughout the water column, predominantly at high [O~2~]. Complexes from clade LLII/III (ex. SS120) and clade LLIV (ex. MIT9313) were detected throughout the water column in high [O~2~] and at depth in low [O~2~] samples. The exclusive assignment of ATP Synthase peptides to Clade IV (ex. MIT9313) may result from the high conservation of protein sequence for this complex.

@@

CladeProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus* clades: HLI (includes *P. marinus* MED4), HLII, LLI, LLII/III (includes P. marinus* SS120) and LLIV (includes *P. marinus* MIT9313).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

*P. marinus* SS120 clade LLII/III,  growth under 250 µM O~2~ increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and across all spectral wavebands (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^ but showed growth inhibition at 16 h red light photoperiod.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white LED, red or blue light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

@@

SS120PAR_cap <- glue("**chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 (Low-Light (LLII/III) deep ocean clade) vs. photoperiod (h). **",
                     "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "colours represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

@@

### SS120 clade LLII/III {.unnumbered}

The representative of the LLII/III clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments (Figure \@ref(fig:PhotoperiodPurFitsPlots)B). Most 4 h photoperiod treatments of SS120 also did not grow under 250 µM O~2~, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments (Figure \@ref(fig:PhotoperiodPurFitsPlots)B). SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments (Figure \@ref(fig:PhotoperiodPurFitsPlots)B).

@@

\newpage
XXXXEdit text depending upon which ocean protein figures we useXXXXX
Ocean detections of proteins mediating protein metabolism support this interpretation of distinct FtsH function across clades of *P. marinus*. Ribosome proteins from clade HLI MED4, clade LLII/III SS120 and clade LLIV MIT9313 show generally similar patterns vs. [O~2~] and depth, a proxy for peak PAR  (Figure \@ref(fig:NATLProchloroProteinMetabDepthO2)).  FtsH3, inferred to mediate PSI assembly, likewise shows a similar pattern between MED4 and MIT9313 (Figure \@ref(fig:NATLProchloroProteinMetabDepthO2)).  But only MED4 shows detected presence of the FtsH1 & FtsH2 isoforms inferred to mediate PSII repair, and then only in near-surface samples subject to higher light.  Furthermore, even though MIT9313 grows (Figure \@ref(fig:MIT9313GAM)), and is detected in the ocean at low [O~2~] (Figure \@ref(fig:CladeProchloroPhotosynthDepthO2)), no FtsH from MIT9313 is detected at low [O~2~] (Figure \@ref(fig:NATLProchloroProteinMetabDepthO2)), suggesting limited requirement for protein turnover under low [O~2~].

@@

NATLProchloroProteinMetabDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* protein metabolism complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (clade HLI), SS120 (clade LLII/III) and MIT9313 (clade LLIV).",
                       "Columns show detections of proteins annotated as FtsH Protease Complexes (FtsH1, FtsH2, FtsH3) or the Ribosome).",
                       "Culture growth experimental conditions indicated by horizontal grey lines for depths approximating Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^) and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

Figure \@ref(fig:ProchlorococcusEnzymeKms) shows the measured or inferred K~M~ for [O~2~] for genes encoding enzymes [@omarAnnotationGenesEncoding2023] for  *P. marinus* strains from clades HLI, LLI, LLII/III and LLIV. XXXX CITATION  to Table once added XXXX. MED4 increases expression of alternative oxidase ('ubiquinol oxidase (non electrogenic)')  to cope with changes in light [@bergResponsesPsbAHli2011], by dissipating electrons from the inter-system transport chain. The approximate K~M~ for [O~2~] of ~ 25 µM for ubiquinol oxidase (non electrogenic) (Figure \@ref(fig:ProchlorococcusEnzymeKms)) is comparable to the lower limit for growth of MED4 in our experiments (Figure \@ref(fig:MED4GAM)).  We suggest that dependence upon this enzyme excludes MED4 from low oxygen zones.  The genome scan shows SS120 and MIT9313 lack this gene (Figure \@ref(fig:ProchlorococcusEnzymeKms)), and therefore, lack this oxygen-dependent path to cope with changing excitation.  Conversely, a gene encoding (S)-2-hydroxy-acid oxidase is encoded in the MIT9313 genome (Figure \@ref(fig:ProchlorococcusEnzymeKms)).  (S)-2-hydroxy-acid oxidase catalyzes the reaction of 2-hydroxy acid with O~2~ to produce toxic H~2~O~2~ [@cunaneCrystalStructureAnalysis2005]. (S)-2-hydroxy-acid oxidase has an approximate K~M~ for [O~2~] of ~ 250 µM, and produces H~2~O~2~, so growth at lower [O~2~] may protect MIT9313 from auto-intoxication from production of H~2~O~2~. We hypothesize that under 250 µM O~2~ and higher blue light, *P. marinus* MIT9313 suffered photoinhibition, resulting from the inactivation of PSII caused by the production of the reactive oxygen species, hydrogen peroxide. This photoinhibition is compounded by the limited inducible repair mechanism for PSII, due to the absence of FtsH 1 and 2 XXXXCHECK TERMINOLOGYXXXX expression in *P. marinus* MIT9313 [@bonisteelStrainSpecificDifferences2018].  We hypothesize that under the conditions of our high light and 2.5 µM or 25 µM O~2~ experiments, the activity of the (S)-2-hydroxy-acid oxidase enzyme is suppressed. As a result, the catalyzed production of hydrogen peroxide is inhibited, leading to less PSII damage, allowing MIT9313 to avoid photoinhibition and circumvent its limitations on PSII repair to exploit higher light.  Figure \@ref(fig:ProchlorococcusEnzymeKms) also shows that *P. marinus* SS120 is the only tested ecotype to lack the pyridoxal 5'-phosphate synthase enzyme. The pyridoxal 5'-phosphate synthase enzyme is an important cofactor in the biosynthesis of vitamin B~6~ [@sangIdentificationPyridoxinePyridoxamine2007]. Vitamin B~6~ is a potential antioxidant and can effectively quench singlet oxygen [@bilskiVitaminB6Pyridoxine2000]. The absence of the pyridoxal 5'-phosphate synthase enzyme may explain why *P. marinus* SS120 does not grow as well as *P. marinus* MIT9313, when exposed to high light stress under 25 µM O~2~ and not at all under 2.5 µM O~2~ (Figure \@ref(fig:SS120PAR)).

@@

DNARepairFig_cap <- glue("**Genes encoding DNA repair enzymes.**" ,
                         "The y-axis represents  *Prochlorococcus marinus* strains.",
                         "The x-axis represents enzymes encoded for DNA repair found in at least one *Prochlorococcus marinus* strain in this study.",
                         "Point size indicate gene counts.",
                       "Figure was generated using a filtered subset of the annotated phytoplankton gene sequences dataset from Omar *et al*. [@omarAnnotationGenesEncoding2023].",

@@

We analyzed  growth rates to determine the viability of *P. marinus* MED4, a clade HLI ecotype found at depths near the ocean surface,
*P. marinus* SS120, a clade LLII/III ecotype found deep in the water column, and *P. marinus* MIT9313, a clade LLIV ecotype also found in deep oceans, under present day and hypothetical future niche conditions. We designed our study to account for light attenuation through ocean waters, recognizing that low light levels of the blue spectral waveband ultimately prevail in deep ocean waters [@morelAvailableUsableStored1978; @holtropVibrationalModesWater2021] and that the solar incidence angle also attenuates the photoperiod or hours of light exposure beneath the ocean surface. We also tested the growth response of *P. marinus* to varying [O~2~] under the matrix of light parameters to determine if *P. marinus* strains can cope with  diverse  ecosystem scenarios.

@@

*Prochlorococcus marinus* SS120, a LLII/III clade representative, showed an interactive inhibition of growth by oxygen and cumulative diel PUR, with a higher tolerance for higher cumulative diel PUR under 25 µM O~2~, compared to 250 µM O~2~. Thus, SS120 can exploit habitats with O~2~ levels spanning 25 µM O~2~ to 250 µM O~2~, including high PAR environments within OMZ.  SS120 is likely excluded from higher [O~2~] and higher PAR because of genomic limitations on capacity for DNA repair (Figure \@ref(fig:DNARepairFig)), and possibly by limited capacity for synthesis of reactive oxygen quenchers (Figure  \@ref(fig:ProchlorococcusEnzymeKms)).  Our growth results are supported by Lavin *et al*. [@lavinNovelLineagesProchlorococcus2010] who found evidence of LLII/III
and LLIV ecotypes, using terminal restriction fragment length polymorphism analyses, at depths above 40 m, where light levels are
higher, within OMZ, and by Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) showing PSII protein subunits annotated as derived from SS120 at all depths ranging
from 20 to 200 m and all [O~2~] in an OMZ of the tropical North Pacific Ocean. SS120, grew under photoperiods longer than 4 h and showed increasing growth rate with increasing photoperiods. Therefore, SS120 can potentially exploit the longer summer photoperiods in temperate zones however, like MED4, cannot sustain growth in winter through its requirement for more than 4 h of light per day, even if water temperature warms into the clade LLII/III tolerance range.

*Prochlorococcus marinus* SS120 prefers the deeper ocean where blue spectral waveband dominates at low levels, through its demonstrated ability to exploit only the lowest blue spectral waveband levels PAR of 30 µmol photons m^-2^ s^-1^ under 250 µM O~2~. However, we found *P. marinus* SS120 can exploit more diverse ecological niches within the ocean layers, even in regions with high levels of blue spectral waveband,  but only under O~2~ of 25 µM (Figure \@ref(fig:SS120PAR)). Lavin *et al*. [@lavinNovelLineagesProchlorococcus2010] show evidence of LLII/III ecotypes at varying depths from 20 to 30 m and from 75 to 200 µM O~2~ in the oxygen minimum zone (OMZ) of the tropical South Pacific Ocean, illustrating its tolerance of high light levels under low O~2~. SS120 has the potential to thrive in deep temperate zones, specifically during the spring, summer, and fall seasons when the duration of daylight exceeds 4 h, if O~2~ are near surface saturation of about 250 µM. Under lower oxygen levels of 25 µM, SS120 can also exploit a 4 h photoperiod in the blue waveband, and thus has the potential to inhabit a warmed, deep, temperate region during the winter season.

@@

PhotoperiodPurFitsPlots_cap <- glue("**Chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                    "Columns are strains; MED4, SS120 and MIT9313.",
                    "Shapes show the imposed photoperiod (h); 4 h (solid square),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; white LED (black symbols), 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Harrison and Platt [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] 4 parameter model fit to data pooled for each combination of strain and dissolved oxygen shown with solid lines. Separate models fit to photoperiod data and shown if significantly different from the pooled model using one-way ANOVA; 4 h (long dashed line); 8 h (dotted line); 12 h (dashed line); and 16 h (dot dashed line).",
                       .sep = " ")

```{r ProchlorococcusLightEnzymes, fig.cap = ProchlorococcusLightEnzymes_cap, fig.height = 10, fig.width = 10, echo = FALSE}
